---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r initial, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  warning = FALSE
)
```

# Transparency & Equity in Property Assessments

This package analyzes the accuracy of property assessments and produces graphs, tables, and reports designed for general use. This package is produced by the [Center for Municipal Finance](https://harris.uchicago.edu/research-impact/centers-institutes/center-municipal-finance), a research institute at the Harris School of Public Policy, University of Chicago.

For a detailed guide on how to use this package check out our [**Get started page**](https://cmf-uchicago.github.io/cmfproperty/articles/cmfproperty.html).

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("cmf-uchicago/cmfproperty")
```

## Basic Usage

First import `cmfproperty`.

```{r import}
library(cmfproperty)
```

To conduct our study, we need data where every roll is a unique property's sale price and assessed value for a given year.

```{r}
head(cmfproperty::example_data)
```


Then, preprocess your data with `reformat_data` and call `make_report`. The report from the example below can be found [here](https://cmf-uchicago.github.io/Cook County, Illinois.html).

```{r reportexample, eval=FALSE}
df <- cmfproperty::example_data

ratios <-
  cmfproperty::reformat_data(
    df,
    sale_col = "SALE_PRICE",
    assessment_col = "ASSESSED_VALUE",
    sale_year_col = "SALE_YEAR",
  )

cmfproperty::make_report(ratios, 
                         jurisdiction_name = "Cook County, Illinois",
                         output_dir = "~/../Documents/GitHub/cmf-uchicago.github.io/") 

#output_dir is the directory in which report is saved; default is working directory

```


