---
title: "Getting Started"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{cmfproperty}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(cmfproperty)
```

# What and Who is `cmfproperty` for?

The property tax is the single largest source of locally generated revenue for American local governments. Cities, counties, school districts, and special districts raise roughly $500 billion per year in property taxes according to [The Urban Institute](https://www.urban.org/policy-centers/cross-center-initiatives/state-and-local-finance-initiative/projects/state-and-local-backgrounders/property-taxes).

Property taxes are supposed to be a *flat tax* where everyone pays the same rate regardless of how much their home is worth. In practice, property taxes are regressive in almost every single county in America, according to a [recent nationwide analysis](https://harris.uchicago.edu/research-impact/centers-institutes/center-municipal-finance) by the Center for Municipal Finance which analyzed 40 million property sales. For an individual homeowner, understanding if their home was assessed fairly is relatively difficult. A homeowner only knows what their home is assessed for and only knows how *accurately* their home was assessed if it sells in a given year.

`cmfproperty` provides the necessary tools for the public, assessors, and public interest groups to evaluate the effectiveness of assessment systems by removing subject-matter and technical barriers and producing automated contextualized reports. These reports analyze the assessor's own data using industry standards and visualizations designed to illustrate inequities **and** package the information in an easy to use format.

# How do I use `cmfproperty`?

First, check out the [article on preprocessing](Preprocessing.html) which outlines data requirements and examples. In order to analyze assessments, data is required to have at least three columns: Sale Year, Sale Price, and Assessed Value.

```{r}
head(cmfproperty::example_data)
```

After gathering data in this form, we have to identify the three required columns (sales, assessments, and sale year) for the package using `cmfproperty::reformat_data`.

```{r}
ratios <-
  cmfproperty::reformat_data(
    data = cmfproperty::example_data,
    sale_col = "SALE_PRICE",
    assessment_col = "ASSESSED_VALUE",
    sale_year_col = "SALE_YEAR",
  )
head(as.data.frame(ratios))
```

We denote data processed by `cmfproperty::reformat_data` as `ratios` throughout the documentation. `ratios` includes the additional calculated fields needed to complete the study.

There's one more step to get the automated report:

```{r eval=FALSE}
cmfproperty::make_report(ratios, 
                         jurisdiction_name = "Cook County, Illinois")
```

The report will be saved in your working directory or a specific directory can be provided.

```{r eval=FALSE}
cmfproperty::make_report(ratios, 
                         jurisdiction_name = "Cook County, Illinois",
                         output_dir = "~/Desktop/")
```

Check out an example report [here](https://cmf-uchicago.github.io/Cook%20County,%20Illinois.html).

# How does this work?

Check out our [methods](Methods.html) and information on [figures](Figures.html) in addition to function-level [documentation](https://cmf-uchicago.github.io/cmfproperty/reference/index.html). The core unit of analysis that we use (and is standard) is the Sales Ratio. A sales ratio is a calculated ratio between the sale price and assessed value at the time of sale for an individual property. Properties should be assessed at their **fair market value** which is defined as the sale price when sold between two unrelated parties in an **arm's length transaction**. We then apply the standard method to determine if a sale is arm's length from the International Association of Assessing Officers.

If you would like to use our package to make your own report, simply take `ratios` and create `stats`.

```{r}
stats <- cmfproperty::calc_iaao_stats(ratios)
head(stats)
```

With `stats` and `ratios`, you have the necessary tools for all the forms of analysis we present. For example,

```{r}
output <- diagnostic_plots(stats, 
                           ratios, 
                           min_reporting_yr = 2015, 
                           max_reporting_yr = 2019)

output[[1]]
```

# Report Diagnostic

We also present a series of diagnostic graphs to evaluate data quality, especially unexpected year-to-year changes. The diagnostic section  can optionally be included as an appendix to the report as follows.

```{r eval=FALSE}
cmfproperty::make_report(ratios, 
                         jurisdiction_name = "Cook County, Illinois Diagnostic",
                         output_dir = "~/../Documents/GitHub/cmf-uchicago.github.io/",
                         diagnostic = TRUE)
```

Check out a report on the example data with a diagnostic section [here](https://cmf-uchicago.github.io/Cook%20County%2C%20Illinois%20Diagnostic.html#5_data_diagnostics).
