---
title: "cmfproperty"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{cmfproperty}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval=FALSE,
  comment = "#>"
)
```

```{r setup}
library(cmfproperty)
```


# Ratios

In this documentation, data which has been appropriately processed is referred to as `ratios` 


`reformat_data` also add the additional calculated fields needed to complete the study:

* RATIO, which is the Sales Ratio (Sale Price / Assessed Value)
* arms_length_transaction, an indicator that the property was sold in an arm's length transaction (calculated using the IAAO standard)
* SALE_PRICE_ADJ, inflation adjusted sale price (adjusted to the last year of available data)
* ASSESSED_VALUE_ADJ, inflation adjusted assessed value (adjusted to the last year of available data)


Note: `ratios` refers to data which has been processed by `reformat_data`



# Other features

## Calculate Regressivity Statistics

This is the basic framework to conduct a sales ratio study:

```{r example}
df <- cmfproperty::example_data
ratios <-
  cmfproperty::reformat_data(
    df,
    sale_col = "SALE_PRICE",
    assessment_col = "ASSESSED_VALUE",
    sale_year_col = "SALE_YEAR",
  )
stats <- cmfproperty::calc_iaao_stats(ratios)
head(stats)
```

## Visualize Regressivity Statistics

```{r example1}
iaao_rslt <- iaao_graphs(stats, ratios, min_reporting_yr = 2015, max_reporting_yr = 2019, "Cook County, Illinois")
```

```{r codgraph}
iaao_rslt[[2]]
```
  
```{r prdgraph}
iaao_rslt[[4]]
```

```{r prbgraph}
iaao_rslt[[6]]
```

## Advanced Regressivity Statistics

```{r example2}
cmfproperty::regression_tests(ratios)
```

## Regressivity Plots

```{r example3}
plot_ls <-
  cmfproperty::plots(ratios,
                     min_reporting_yr = 2015,
                     max_reporting_yr = 2019,
                     jurisdiction_name = "Cook County, Illinois")

```

`r #plot_ls[[1]]`

```{r plot1}
plot_ls[[2]]
```

`r #plot_ls[[3]]`

```{r plot2}
plot_ls[[4]]
```

## Monte Carlo Analysis

```{r montecarlo, fig.height=5}
m_rslts <- monte_carlo_graphs(ratios)
gridExtra::grid.arrange(m_rslts[[1]], m_rslts[[2]], m_rslts[[3]], m_rslts[[4]], m_rslts[[5]], m_rslts[[6]], nrow = 3)
```

## Report Evaluation Output
```{r evaluatedreport, eval=FALSE}
df <- cmfproperty::example_data

ratios <-
  cmfproperty::reformat_data(
    df,
    sale_col = "SALE_PRICE",
    assessment_col = "ASSESSED_VALUE",
    sale_year_col = "SALE_YEAR",
  )

cmfproperty::make_report(ratios, 
                         jurisdiction_name = "Cook County, Illinois",
                         output_dir = "C:/Users/erhla/Desktop/")
```
