---
title: "Figures"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Figures}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, results='hide', message=FALSE, echo=FALSE, warning = FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE)
library(cmfproperty)
library(knitr)
library(kableExtra)
library(rmarkdown)
library(data.table)

ratios <-
  cmfproperty::reformat_data(
    data = cmfproperty::example_data,
    sale_col = "SALE_PRICE",
    assessment_col = "ASSESSED_VALUE",
    sale_year_col = "SALE_YEAR",
  )
stats <- cmfproperty::calc_iaao_stats(ratios)
```

This article produces a gallery of figures and tables produced by this package for reference.

# calc_iaao_stats
``` {r iaao_stats, echo=FALSE}
stats_colnames_copy <- copy(stats)
names(stats_colnames_copy) <- tolower(names(stats))
paged_table(stats_colnames_copy)
```

# regression_tests
``` {r regression_tests, fig.width=6, echo=FALSE}
models <- cmfproperty::regression_tests(ratios, TRUE, TRUE)
plots <- diagnostic_plots(stats, ratios, min_reporting_yr = 2015, max_reporting_yr = 2019)
plots[[1]]
plots[[2]]
plots[[3]]
plots[[4]]
plots[[5]]
gridExtra::grid.arrange(plots[[6]], plots[[7]], plots[[8]], plots[[9]], ncol=2, nrow=2)
```


# iaao_graphs
```{r example1, message=FALSE, echo=FALSE}
iaao_rslt <- cmfproperty::iaao_graphs(stats, ratios, min_reporting_yr = 2015, max_reporting_yr = 2019, "Cook County, Illinois")
```

## Coefficient of Dispersion (COD)
```{r codgraph, echo=FALSE, fig.width=6, fig.height=3}
iaao_rslt[[2]]
```

## Price-Related Differential (PRD)
```{r prdgraph, echo=FALSE, fig.width=6, fig.height=3}
iaao_rslt[[4]]
```

## Coefficient of Price-Related Bias (PRB)
```{r prbgraph, echo=FALSE, fig.width=6, fig.height=3}
iaao_rslt[[6]]
```


# monte_carlo_graphs
```{r montecarlo_graphs, echo=FALSE, fig.width=6, fig.height=5}
m_rslts <- cmfproperty::monte_carlo_graphs(ratios)
gridExtra::grid.arrange(m_rslts[[1]], m_rslts[[2]], m_rslts[[3]], m_rslts[[4]], m_rslts[[5]], m_rslts[[6]], nrow = 3)
```

# diagnostic_plots
``` {r diagnostic, echo=FALSE, fig.width=6, fig.height=4}
plots <- diagnostic_plots(stats, ratios, min_reporting_yr = 2015, max_reporting_yr = 2019)
plots[[1]]
plots[[2]]
plots[[3]]
plots[[4]]
```

# binned_scatter
``` {r binned, echo=FALSE, fig.width=7, fig.height=4}
binned <- cmfproperty::binned_scatter(ratios, 
                                      min_reporting_yr = 2015,
                                      max_reporting_yr = 2019,
                                      params$title)
binned[[2]]
```


# pct_over_under
``` {r pct_over}
# pct_over <- cmfproperty::pct_over_under(ratios,
#                                         min_reporting_yr = 2015,
#                                         max_reporting_yr = 2019,
#                                         params$title)
```





## Advanced Regressivity Statistics

```{r example2, echo=FALSE}
paged_table(cmfproperty::regression_tests(ratios))
```

## Regressivity Plots

```{r example3}
# plot_ls <-
#   cmfproperty::plots(ratios,
#                      min_reporting_yr = 2015,
#                      max_reporting_yr = 2019,
#                      jurisdiction_name = "Cook County, Illinois")
# update to use binned_scatter and pct_over_under
```

## Monte Carlo Analysis

```{r montecarlo, fig.width=5, fig.height=5}
# m_rslts <- monte_carlo_graphs(ratios)
# gridExtra::grid.arrange(m_rslts[[1]], m_rslts[[2]], m_rslts[[3]], m_rslts[[4]], m_rslts[[5]], m_rslts[[6]], nrow = 3)
```













