---
title: "Methods"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Methods}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(cmfproperty)
```

This article is a reference section to verify that this package calculates various metrics correctly.

# IAAO Functions

## Coefficient of Dispersion (COD)

$$100 * \frac{\sum_{i=1}^N(SalesRatio_i - SalesRatio_{Median})}{N * SalesRatio_{Median}}$$

## Price-Related Differential (PRD)

$\frac{\frac{\sum_{n=1}^NSalesRatio_i}{N}} {\frac{\sum_{n=1}^N(SalesRatio_i*SalePrice_i)} {N}}$

## Coefficient of Price-Related Bias (PRB)
## Arms-Length Measure

# Other Measures

## From paper X

### Paglin

$$ AssessedValue = \alpha_0 + \alpha_1SalesPrice + \epsilon $$
### Cheng

$$ ln(AssessedValue) = \alpha_0 + \alpha_1ln(SalesPrice) + \epsilon $$
### IAAO

$$ AssessedRatio = \alpha_0 + \alpha_1SalesPrice + \epsilon $$
### Kochin

$$ ln(SalesPrice) = \alpha_0 + \alpha_1ln(Assessed Value) + \epsilon $$
### Bell

$$ AssessedValue = \alpha_0 + \alpha_1SP + \alpha_2SP^2 + \epsilon$$
### Sunderman

$$ AssessedValue = \alpha_{00} + \alpha_{10}SalesPrice + \alpha_{01}Low + \alpha_{02}High + \alpha_{11}LowSalesPrice + \alpha_{12}HighSalesPrice + \epsilon $$

## Inflation Adjustment
